Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-17 paper.qmd
Try to update with new function

  a1_1 <- hb_obj$a1

@@

  return(list(nldr_scaled = nldr_scaled,
              hex_grid_with_counts = hex_grid_with_counts,
              hex_grid_nonempty = hex_grid_nonempty,
              df_bin_centroids = df_bin_centroids,
              tr_from_to_df = tr_from_to_df,
              df_bin = df_bin,
              distance_df = distance_df,
              benchmark = benchmark))

@@

<!--tSNE applied for Two curvy data-->
```{r}
#| label: tsne-two-curvy
tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")

two_curvy_scaled_obj <- gen_scaled_data(
  data = tsne_two_curvy)

tsne_two_curvy_scaled <- two_curvy_scaled_obj$scaled_nldr |>
  mutate(ID = row_number())

lim1 <- two_curvy_scaled_obj$lim1
lim2 <- two_curvy_scaled_obj$lim2
r2 <- diff(lim2)/diff(lim1)

sc_ltr_pos <- c(0.08, 0.9)
sc_xlims <- c(-0.3, 1.2)
sc_ylims <- c(-0.17, 1.25)

@@

# Third plot with aspect ratio and different annotation
nldr_two_curvy_original2 <- nldr_two_curvy_original +
  theme(aspect.ratio = 1) +
  interior_annotation("a1", sc_ltr_pos)
```

```{r}
#| label: hexbin-two-curvy
## Compute hexbin parameters
num_bins_x_two_curvy <- 22

## hexagon binning to have regular hexagons
hb_obj_two_curvy <- hex_binning(
  data = tsne_two_curvy_scaled, 
  bin1 = num_bins_x_two_curvy, 
  r2 = r2,
  q = 0.1)

## Data set with all centroids
all_centroids_df <- hb_obj_two_curvy$centroids

a1_1 <- hb_obj_two_curvy$a1

## Generate all coordinates of hexagons
hex_grid <- hb_obj_two_curvy$hex_poly

## To obtain the standardise counts within hexbins
counts_df <- hb_obj_two_curvy$std_cts
df_bin_centroids_two_curvy <- extract_hexbin_centroids(
  centroids_df = all_centroids_df, 
  counts_df = counts_df) |>
  filter(drop_empty == FALSE)

umap_data_two_curvy_with_hb_id <- hb_obj_two_curvy$data_hb_id
df_all_two_curvy <- dplyr::bind_cols(training_data_two_curvy, umap_data_two_curvy_with_hb_id)
df_bin_two_curvy <- avg_highd_data(data = df_all_two_curvy, col_start = "x")

@@

```{r}
#| label: empty-bin-two-curvy
hex_grid_nonempty <- hex_grid |>
  filter(hex_poly_id %in% df_bin_centroids_two_curvy$hexID)

@@

<!--2D model-->
```{r}
#| label: triangulate-two-curvy
## Triangulate bin centroids
tr1_object_two_curvy <- tri_bin_centroids(
  df_bin_centroids_two_curvy, x = "c_x", y = "c_y")
tr_from_to_df_two_curvy <- gen_edges(
  tri_object = tr1_object_two_curvy)

@@

## Compute 2D distances
distance_two_curvy <- cal_2d_dist(
  tr_coord_df = tr_from_to_df_two_curvy, 
  start_x = "x_from", 
  start_y = "y_from", 
  end_x = "x_to", 
  end_y = "y_to", 
  select_vars = c("from", "to", "distance"))

@@

bin_width <- hb_obj_two_curvy$a1

####With diff benchmark values

## Benchmark 2

benchmark_two_curvy <- 2.8 * bin_width

distance_df_small_edges_two_curvy3 <- distance_two_curvy |>
  filter(distance < benchmark_two_curvy)

tr_from_to_df_two_curvy4 <- inner_join(
  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy3,
  by = c("from", "to"))

trimesh_two_curvy_removed2 <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy4,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000")+
  geom_point(data = tsne_two_curvy_scaled,
             aes(
               x = tSNE1,
               y = tSNE2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5
  ) +
  coord_equal() +
  interior_annotation("b1", sc_ltr_pos)

benchmark_two_curvy <- 2 * bin_width

distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
  filter(distance < benchmark_two_curvy) ## 0.231

tr_from_to_df_two_curvy3 <- inner_join(
  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy2,
  by = c("from", "to"))

trimesh_two_curvy_removed1 <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy3,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000")+
  geom_point(data = tsne_two_curvy_scaled,
             aes(
               x = tSNE1,
               y = tSNE2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5
  ) +
  coord_equal() +
  interior_annotation("c1", sc_ltr_pos)

trimesh_two_curvy_removed1_with_data <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy3,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = tsne_two_curvy_scaled,
             aes(
               x = tSNE1,
               y = tSNE2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5
  ) +
  coord_equal() +
  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
  interior_annotation("a2", sc_ltr_pos)

>>>>

  a1 <- hb_obj$a1

@@

  return(list(nldr_scaled = nldr_scaled,
              a1 = a1,
              hex_grid_with_counts = hex_grid_with_counts,
              hex_grid_nonempty = hex_grid_nonempty,
              df_bin_centroids = df_bin_centroids,
              tr_from_to_df = tr_from_to_df,
              df_bin = df_bin,
              distance_df = distance_df,
              benchmark = benchmark))

@@

<!--tSNE applied for Two curvy data-->
```{r}
#| label: tsne-two-curvy
tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds") 
```

```{r}
#| label: fit_model-two-curvy

num_bins_x_two_curvy <- 22

algo_obj_two_curvy <- gen_nldr_vis_algo_obj(
  high_d_data = training_data_two_curvy, 
  nldr_data = tsne_two_curvy, 
  num_x_bins = num_bins_x_two_curvy,
  col_start = "x")

tsne_two_curvy_scaled <- algo_obj_two_curvy$nldr_scaled
distance_two_curvy <- algo_obj_two_curvy$distance_df
tr_from_to_df_two_curvy <- algo_obj_two_curvy$tr_from_to_df
benchmark_two_curvy <- algo_obj_two_curvy$benchmark
df_bin_centroids_two_curvy <- algo_obj_two_curvy$df_bin_centroids
df_bin_two_curvy <- algo_obj_two_curvy$df_bin
hex_grid_with_counts <- algo_obj_two_curvy$hex_grid_with_counts
hex_grid_nonempty <- algo_obj_two_curvy$hex_grid_nonempty
bin_width <- algo_obj_two_curvy$a1

sc_ltr_pos <- c(0.08, 0.9)
sc_xlims <- c(-0.3, 1.2)
sc_ylims <- c(-0.17, 1.25)
```

```{r}
#| label: hexbin-two-curvy

@@

# Third plot with aspect ratio and different annotation
nldr_two_curvy_original2 <- nldr_two_curvy_original +
  theme(aspect.ratio = 1) +
  interior_annotation("a1", sc_ltr_pos)

@@

```{r}
#| label: empty-bin-two-curvy

@@

<!--2D model-->
```{r}
#| label: triangulate-two-curvy

@@

@@
